---
title: "Data Visualization Exercises: 1"
author: "Kieran Healy"
date: "8/3/2020"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Overview 

In these exercises, you will use R, RStudio, and ggplot to look at the `gapminder` data again. Our goal is to create an RStudio project, load the required libraries, and begin looking at our data. 

## 1  Set Up a Project

- **Create** a new RStudio project and place it on your computer somewhere. Call it something like "Exercises"
- Open that new folder in Windows File Explorer or macOS Finder (however you navigate around the files on your computer), and create a folder there named `figures` 
- **Create** a new R Markdown file and save it in your project. 
- In RStudio go to File > New File > R Markdown…, choose the default options, and delete all the placeholder text in the new file except for the metadata at the top, which is between `---` and `---`. 
- **Verify** that your project folder is structured like this:

    ```
    your-project-name/
      your-analysis.Rmd
      your-project-name.Rproj
      figures/
        <EMPTY>
    ```

## 2 Get the data

- **Load** the libraries required for the analysis. Create an R code chunk that runs this code:

```{r libraries}

library(tidyverse)
library(gapminder)
library(socviz)

```

## 3 Examine the Gapminder Data

```{r gapminder}
gapminder
```


- **Create** the following plots (Make a new code chunk for each plot)

1. A scatter plot (using `geom_point()`) of GDP per Capita (on the x-axis) and Life Expectancy (on the y-axis). 
2. Adjust the plot so that the x-axis scale is in log units rather than raw GDP, and properly label the axes. 
3. Using `subset()` function, choose a *single year* and make a histogram of country populations. *Hint:* Use `==` and not `=` when asking `subset` to pick a particular value of `year`.
5. Try adding a log scale to the x-axis.
6. Facet the plot by continent.
7. Subset the data so that you pick out the following four years: 1952, 1972, 1992, and 2002. Plot a histogram of life expectancy faceted by year. 

## 4 Further Questions and Topics

- What happens when you put the `geom_smooth()` function before `geom_point()` instead of after it?
- Plot life expectancy against _population_ rather than per capita GDP. What does that look like? 
- Try some alternative scale mappings beyond `scale_x_log10()`, such as `scale_x_sqrt()` and `scale_x_reverse()`. There are corresponding functions for y-axis transformations. 
- What happens if you map `color` to `year` instead of `continent`? Is the result what you expected?
- Is the example figure we worked up and polished during class really the best way to display the country-year data in `gapminder`? How might we do better?

## 5 Produce your answers as an HTML, PDF, or Word document

- Once you've written your code, knit your document.
