---
title: "Data Visualization Exercises: 2"
author: "Kieran Healy"
date: "8/3/2020"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1 Overview 

Working within the project you created, you'll explore the GSS data in `gss_sm` a little more.

## 2 Setup

- **Load** the libraries required for the analysis. Create an R code chunk that runs this code:

```{r libraries}

library(tidyverse)
library(socviz)

```

## 3 Examine the GSS data

```{r gss}
gss_sm
```

## 4 Tables and Graphs

### Example

```{r example}
gss_sm %>% 
    group_by(degree, kids) %>%
    tally() %>%
    drop_na() %>% ## This function drops any row in your table with missing values
    mutate(pct = n/sum(n) * 100) %>%
    ggplot(mapping = aes(x = kids, y = pct)) +
    geom_col() +
    facet_wrap(~ degree)
```

### Things to try 

- Make the facet so that there is only one row in the plot. Then try it with only one column.
- Create the plot above, but for respondents from the Northeast only
- Add labels and a title


- The GSS asks respondents how happy they are feeling. In `gss_sm` it's the `happy` column. Make a graph of the measure. 
- Make a table and then a graph of happiness by sex of respondent.
- Try the same with happiness and marital status.
- What graph options are available for a measure like this? Which seem best to you? Experiment with `geom_col()`, `geom_point()`, and `geom_line()`. 
- Explore graphing the relationship between happiness and age. Options include choosing to focus on just one of the happiness measures. Can `geom_boxplot()` be of any use?

- Practice summarizing at different levels of grouping. Can you produce, for example, two- or three-way summaries of average number of children by sex and degree, or sex, degree, and partyid? Can you use `geom_point()` to produce faceted dotplots of the results? (Hint 1: remember to pass `na.rm =  TRUE` when invoking `mean()`. Hint 2: try reordering the facets by the variable you're summarizing.) 


## 5 Produce your answers as an HTML, PDF, or Word document

- Once you've written your code, knit your document.
